---
title: "Canopy Structure Manuscript RF Predictions"
author: "Roí Ankori-Karlinsky"
date: "3/7/2023"
output:
  html_document: 
    toc: yes
    toc_depth: 5
    toc_float: yes
    theme: journal
    fig_width: 9
    fig_height: 8
---


```{r setup, include=FALSE,echo=FALSE}

# load packages 

#Basic
library(tidyverse) # Grammar for data manipulation
library(parallel) # mclapply for multicore processing
library(vegan)

#Spatial analysis
library(sp) #provides object oriented classes for spatial data in R 
library(sf) #similar to sp
library(rgdal) # raster analysis
library(raster) #provides classes and methods for raster datasets
library(spdplyr) #spatial data manipulation

#Data checking
library(naniar) # Visualization of missing values
library(Amelia) # Teting missing columns
library(corrplot) #correlation plots

#Random Forest
library(randomForestSRC) # random forests for survival, regression and
library(pdp) #partial dependence plots
library(randomForest) #random forest
library(randomForestExplainer) #unpacking RF results
library(pre) #Interactions between predictors
library(iml) #Interactions between predictors

# Visualization
library(ggRandomForests) # ggplot2 random forest figures
library(gridExtra) # allows plotting of multiple ggplots in one
library(grid) #same as above
library(colorRamps) # Colors for graphing
library(RColorBrewer) # Colors for graphing
library(cowplot) #putting plots together
library(plot3D) #Surface plots

#Making sure R is utilizing all available RAM
memory.limit(size=50000)

#Don't include code chunks when outputting markdown
knitr::opts_chunk$set(
echo=FALSE
)

```

```{r data prep, message=FALSE, warning=FALSE, echo=F}

#File path
local_path <- 'C:/Users/roiak/Documents/Damage Project/Rugosity'

full_d <- read_csv(paste0(local_path,"/Data/Final/Full_Data_For_Analyses.csv"))#data file

#cleaning data
full_d <- full_d %>%
  filter(Soil!="Water") %>%
  #Fixing variables (factors, making sure precipitation is in mm/year)
  mutate(Soil = factor(Soil),
         Max_Height = as.numeric(Max_Height),
         Forest_Age = factor(Forest_Age),
         MAP = MAP *0.01,
         Wind_Binary = ifelse(Wind_Binary == 0, "Protected",
                              ifelse(Wind_Binary == 1, "Exposed",NA)),
         Wind_Exposure = round(Wind_Exposure,2),
         Wind_sd = round(Wind_sd,2)
         )  %>%
  #Removing outliers and sites below sea-level
  filter(CHM > 2 & CHM <40, #getting rid of canopy height outliers
         Elev_Mean >=0)

#Creating dataset for each response variable
height <- full_d %>%
  dplyr::select(CHM,
                Forest_Age,
                MAP,
                Wind_Exposure,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean,
                Hugo_exposure) %>%
  na.omit(.)


rugosity <- full_d %>%
  dplyr::select(Rugosity,
                CHM,
                Forest_Age,
                MAP,
                Wind_Exposure,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean,
                Hugo_exposure) %>%
  na.omit(.)

```

# OVERVIEW

## Questions

- **(1)** Does regular exposure to non-hurricane winds reduce canopy structural complexity in tropical forests?  
- **(2)** To what extent is this effect mediated by climate, topography, substrate, forest age and past disturbances?

## Goals

This script runs Random Forest (RF) models of two canopy metrics:    
      - **(1)** Canopy Height
      - **(2)** Canopy Top Rugosity (sd of height along moving windows) - Vertical variation for top of canopy  
      
      
It then runs partial dependence plots to see how they're predicted by:  
      - Soil type and soil water availability
      - Forest age  
      - Wind exposure (calculated with prevalent directions and using a 30m DEM)
      - Elevation mean and range (30m USGS DEM)
      - Slope (30m USGS DEM)
      - Long term mean annual precipitation 1963-1995 (PRISM)
      - Past exposure to hurricane Hugo in 1989 and hurricane Georges in 1998 (Boose et al., 2004)

## Data

The LiDAR data come from 2016 flights by the USGS, which were then normalized by Lora Murphy of the Cary Institute.
Lora also calculated a canopy height model (CHM) and top rugosity.

Study sites were selected by stratified sampling randomly across forest age,
Then creating a required 250m minimum distance between points,
Then creating 30m buffers around to have a scale that captures a forest stand,
Then filtering out non-forest points if they contain non-forest elements within the buffer.

This resulted in 20,660 study sites across Puerto Rico for the 30m buffer.

(Thanks Dr. Jazlynn Hall with help for the code!)


***

# Random Forest Models  

For each response variable, I will  
- **(1)** Run and save the RF model
- **(2)** Run partial dependence plots and save the model predictions 
- **(3)** Run partial dependence coplots of each predictor against wind exposure and save predictions  


## Canopy Height 

### Running random forest model

```{r Canopy Height RF, message=FALSE, warning=FALSE}

# Run the random forest model with height as response
# (making sure data is dataframe not tibble)
RandomForest_Model <- rfsrc(CHM~.,
                            data=as.data.frame(height),
                            ntree = 5000, #number of trees
                            importance = "permute") #return VIMP calculated with permutation


# print the forest summary
RandomForest_Model

```

### Partial Dependence Predictions 

```{r Canopy Height PD Predictions, message=FALSE, warning=FALSE}

# Save the gg_minimal_depth object for later use.
gg_md <- gg_minimal_depth(var.select(RandomForest_Model))

# Calculate partial dependence: THIS TAKES A WHILE
Model_partials <- plot.variable(RandomForest_Model,
                            xvar=gg_md$topvars,
                            partial=TRUE, sorted=TRUE,
                            show.plots = FALSE)

# generate a list of gg_partial objects, one per xvar.
gg_p <- gg_partial(Model_partials)

#save predictions
save(gg_p,
     file=paste0(local_path,"/R Code/Height_PD.RData"))

```

### Partial Dependence Coplots

Here I divide wind exposure into 4 probabilities to understand its interaction with other predictors

```{r Canopy Height PD Wind Break, message=FALSE, warning=FALSE}

#Breaking wind exposure into quartiles
wind_pts <- quantile_pts(RandomForest_Model$xvar$Wind_Exposure, groups=4, intervals=TRUE)

# Pass these variable points to create the 4 (factor) intervals
wind_grp <- cut(RandomForest_Model$xvar$Wind_Exposure, 
                breaks=wind_pts)

```


#### Wind and Precipitation Coplots

```{r Canopy Height PD Wind Prec Coplot, message=FALSE, warning=FALSE, fig.width=9}

#Running partial dependence coplot for precipitation effect on height grouped by wind exposure factor
partial_coplot <- gg_partial_coplot(RandomForest_Model,
                                    xvar="MAP",
                                    groups=wind_grp,
                                    show.plots=FALSE)

#saving object
save(partial_coplot,
     file=paste0(local_path,
                 "/R Code/RF_Height_WindPrec.RData"))


```

```{r cleaning, echo=F}
rm(partial_coplot)
gc()
```

### Wind and Elevation Coplots

```{r Canopy Height PD Wind Elev Coplot, message=FALSE, warning=FALSE, fig.width=9}

#Running partial dependence coplot for precipitation effect on height grouped by wind exposure factor
partial_coplot <- gg_partial_coplot(RandomForest_Model,
                                    xvar="Elev_Mean",
                                    groups=wind_grp,
                                    show.plots=FALSE)

#saving object
save(partial_coplot,
     file=paste0(local_path,
                 "/R Code/RF_Height_WindElev.RData"))
```


```{r cleaning2, echo=F}
rm(partial_coplot)
gc()
```

#### Wind and Forest Age Coplots

```{r Canopy Height PD Wind Age Coplot, message=FALSE, warning=FALSE, fig.width=9}

#Run partial depednence coplot
partial_coplot <- gg_partial_coplot(RandomForest_Model,
                                    xvar="Forest_Age",
                                    groups=wind_grp,
                                    show.plots=FALSE)

#fixing levels
partial_coplot$Forest_Age <- factor(partial_coplot$Forest_Age,
                                    levels=c("1","2","3","4","5"))

#saving object
save(partial_coplot,
     file=paste0(local_path,
                 "/R Code/RF_Height_WindAge.RData"))

```


```{r cleaning3}
rm(partial_coplot,RandomForest_Model,gg_p,gg_md)
gc()

```

***

## Canopy Top Rugosity

### Running random forest model for canopy top rugosity

```{r Canopy Rugosity RF, message=FALSE, warning=FALSE}

# Run the random forest model with rugosity as response
# (making sure data is dataframe not tibble)
RandomForest_Model <- rfsrc(Rugosity~.,
                            data=as.data.frame(rugosity),
                            ntree = 5000, #number of trees
                            importance = "permute") #return VIMP calculated with permutation


# print the forest summary
RandomForest_Model

```

### Partial Dependence Predictions 

```{r Canopy Rugosity PD Predictions, message=FALSE, warning=FALSE}

# Save the gg_minimal_depth object for later use.
gg_md <- gg_minimal_depth(var.select(RandomForest_Model))

# Calculate partial dependence: THIS TAKES A WHILE
Model_partials <- plot.variable(RandomForest_Model,
                            xvar=gg_md$topvars,
                            partial=TRUE, sorted=TRUE,
                            show.plots = FALSE)

# generate a list of gg_partial objects, one per xvar.
gg_p <- gg_partial(Model_partials)

#save predictions
save(gg_p,
     file=paste0(local_path,"/R Code/Rugosity_PD.RData"))

```

### Partial Dependence Coplots

Here I divide wind exposure into 4 probabilities to understand its interaction with other predictors

```{r Canopy Rugosity PD Wind Break, message=FALSE, warning=FALSE}

#Breaking wind exposure into quartiles
wind_pts <- quantile_pts(RandomForest_Model$xvar$Wind_Exposure, groups=4, intervals=TRUE)

# Pass these variable points to create the 4 (factor) intervals
wind_grp <- cut(RandomForest_Model$xvar$Wind_Exposure, 
                breaks=wind_pts)

```

#### Wind and Height Coplot

```{r Canopy Rugosity PD Wind Height Coplot, message=FALSE, warning=FALSE, fig.width=9}

#Running partial dependence coplot for height effect on rugosity grouped by wind exposure factor
partial_coplot <- gg_partial_coplot(RandomForest_Model,
                                    xvar="CHM",
                                    groups=wind_grp,
                                    show.plots=FALSE)

#saving object
save(partial_coplot,
     file=paste0(local_path,
                 "/R Code/RF_Rugosity_WindHeight.RData"))

```

```{r cleaning4, echo=F}
rm(partial_coplot)
gc()
```

#### Wind and Precipitation Coplots

```{r Canopy Rugosity PD Wind Prec Coplot, message=FALSE, warning=FALSE, fig.width=9}

#Running partial dependence coplot for precipitation effect on rugosity grouped by wind exposure factor
partial_coplot <- gg_partial_coplot(RandomForest_Model,
                                    xvar="MAP",
                                    groups=wind_grp,
                                    show.plots=FALSE)

#Saving object
save(partial_coplot,
     file=paste0(local_path,
                 "/R Code/RF_Rugosity_WindPrec.RData"))

```

```{r cleaning5, echo=F}
rm(partial_coplot)
gc()
```

#### Wind and Elevation Coplots

```{r Canopy Rugosity PD Wind Elev Coplot, message=FALSE, warning=FALSE, fig.width=9}

#Running partial dependence coplot for precipitation effect on rugosity grouped by wind exposure factor
partial_coplot <- gg_partial_coplot(RandomForest_Model,
                                    xvar="Elev_Mean",
                                    groups=wind_grp,
                                    show.plots=FALSE)

#saving object
save(partial_coplot,
     file=paste0(local_path,
                 "/R Code/RF_Rugosity_WindElev.RData"))

```

```{r cleaning6, echo=F}
rm(partial_coplot)
gc()
```

#### Wind and Forest Age Coplots


```{r Canopy Rugosity PD Wind Age Coplot, message=FALSE, warning=FALSE, fig.width=9}

partial_coplot <- gg_partial_coplot(RandomForest_Model,
                                    xvar="Forest_Age",
                                    groups=wind_grp,
                                    show.plots=FALSE)

#fixing levels
partial_coplot$Forest_Age <- factor(partial_coplot$Forest_Age,
                                    levels=c("1","2","3","4","5"))

#saving object
save(partial_coplot,
     file=paste0(local_path,
                 "/R Code/RF_Rugosity_WindAge.RData"))

```

*** 


```{r cleaning final}
#cleaning
rm(list=ls())
gc()

```