---
title: "Canopy Structure Manuscript Analyses"
author: "Roí Ankori-Karlinsky"
date: "3/7/2023"
output:
  html_document: 
    toc: yes
    toc_depth: 5
    toc_float: yes
    theme: journal
    fig_width: 9
    fig_height: 8
---


```{r setup, include=FALSE,echo=FALSE}

# load packages 

#Basic
library(tidyverse) # Grammar for data manipulation
library(parallel) # mclapply for multicore processing
library(vegan)

#Spatial analysis
library(sp) #provides object oriented classes for spatial data in R 
library(sf) #similar to sp
library(rgdal) # raster analysis
library(raster) #provides classes and methods for raster datasets
library(spdplyr) #spatial data manipulation

#Data checking
library(naniar) # Visualization of missing values
library(Amelia) # Teting missing columns
library(corrplot) #correlation plots

#Random Forest
library(randomForestSRC) # random forests for survival, regression and
library(pdp) #partial dependence plots
library(randomForest) #random forest
library(randomForestExplainer) #unpacking RF results
library(pre) #Interactions between predictors
library(iml) #Interactions between predictors

# Visualization
library(ggRandomForests) # ggplot2 random forest figures
library(gridExtra) # allows plotting of multiple ggplots in one
library(grid) #same as above
library(colorRamps) # Colors for graphing
library(RColorBrewer) # Colors for graphing
library(cowplot) #putting plots together
library(plot3D) #Surface plots

#Making sure R is utilizing all available RAM
memory.limit(size=50000)

#Don't include code chunks when outputting markdown
knitr::opts_chunk$set(
echo=FALSE
)

```

```{r data prep, message=FALSE, warning=FALSE, echo=F}

#File path
local_path <- 'C:/Users/roiak/Documents/Damage Project/Rugosity'

full_d <- read_csv(paste0(local_path,"/Data/Final/Full_Data_For_Analyses.csv"))#data file

#cleaning data
full_d <- full_d %>%
  filter(Soil!="Water") %>%
  #Fixing variables (factors, making sure precipitation is in mm/year)
  mutate(Soil = factor(Soil),
         Max_Height = as.numeric(Max_Height),
         Forest_Age = factor(Forest_Age),
         MAP = MAP *0.01,
         Wind_Binary = factor(Wind_Binary,
                              levels=c(0,1))
         )  %>%
  #Removing outliers and sites below sea-level
  filter(CHM > 2 & CHM <40, #getting rid of canopy height outliers
         Elev_Mean >=0)

#Creating dataset for each response variable
height <- full_d %>%
  dplyr::select(CHM,
                Forest_Age,
                MAP,
                Wind_Binary,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean,
                Hugo_exposure) %>%
  na.omit(.)


rugosity <- full_d %>%
  dplyr::select(Rugosity,
                CHM,
                Forest_Age,
                MAP,
                Wind_Binary,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean,
                Hugo_exposure) %>%
  na.omit(.)

```

# OVERVIEW

### Questions

- **(1)** Does regular exposure to non-hurricane winds reduce canopy structural complexity in tropical forests?  
- **(2)** To what extent is this effect mediated by climate, topography, substrate, forest age and past disturbances?

### Goals

This script tests whether chronic wind exposure reduces canopy structural complexity across Puerto Rico.
It does this by running 2 Random-Forest (RF) models for two airborne LiDAR-derived canopy metrics:

- **(1)** Canopy Height
- **(2)** Canopy Top Rugosity (sd of height along moving windows) - Vertical variation for top of canopy  

### Data

The LiDAR data come from 2016 flights by the USGS, which were then normalized by Lora Murphy of the Cary Institute.
Lora also calculated a canopy height model (CHM) and top rugosity.

Study sites were selected by stratified sampling randomly across forest age,
Then creating a required 250m minimum distance between points,
Then creating 30m buffers around to have a scale that captures a forest stand,
Then filtering out non-forest points if they contain non-forest elements within the buffer.

This resulted in 20,660 study sites across Puerto Rico for the 30m buffer.

The independent variables used are:
  - Soil type and soil water availability
  - Forest age  
  - Wind exposure (calculated with prevalent directions and using a 30m DEM)
  - Elevation mean and range (30m USGS DEM)
  - Slope (30m USGS DEM)
  - Long term mean annual precipitation 1963-1995 (PRISM)
  - Past exposure to hurricane Hugo in 1989 and hurricane Georges in 1998 (Boose et al., 2004)

***

# Random Forest Models

Here I run 10 RF models for each response variable.
Then I summarize variable importance of predictors as the mean and standard error from all 10 models.
I also report the model error and R$^2$ 


## Canopy Height


### Running RF 

```{r Canopy Height RF, message=FALSE, warning=FALSE}

#empty matrix to store error and r2 from model
Height_model<- matrix(nrow=10,ncol=2)

#empty dataframe to store variable importance
Height_VIMP <- data.frame()


#run model 10 times and store results in empty dataframes
for(i in 1:10){

  #----RUNNING MODEL----#
  
  # Run the random forest model with height as response
  # (making sure data is dataframe not tibble)
  RF_Height<- rfsrc(CHM~.,
                            data=as.data.frame(height),
                            ntree = 5000, #number of trees
                            importance = "permute") #return VIMP calculated with permutation
  
  #----STORING RESULTS----#
  
  # R2
  Height_r2_i <- round(100* (1-mean(RF_Height$err.rate,na.rm=T)/var(RF_Height$yvar, na.rm = TRUE)),2)
  # error rate
  Height_error_i <- mean(RF_Height$err.rate,na.rm=T)
  # storing
  Height_model[i,] <- c(Height_r2_i,Height_error_i)
  
  #----STORING VIMP----#
  
  # Get variable importance and organize it
  Height_vimp_i <- as.data.frame(RF_Height$importance) %>%
    rownames_to_column(var = "Variable") %>%
    dplyr::rename(Variable_Importance = `RF_Height$importance`) %>%
    arrange(Variable_Importance)
  
  # storing
  Height_VIMP = rbind(Height_VIMP,Height_vimp_i)

}

```


### Summarizing Results

```{r Height Results Summarize, message=FALSE, warning=FALSE}

# turning matrix of results to dataframe
Height_model <- as.data.frame(Height_model)
colnames(Height_model) <- c("R2","Error")

# summarizing
Height_results <- Height_model %>%
  mutate(Mean_R2 = round(mean(R2,na.rm=T),2)*0.01,
         R2_sd = round(sd(R2,na.rm=T),2)*0.01,
         Mean_OOB_Error = round(mean(Error,na.rm=T),2)*0.01,
         Error_sd = round(sd(Error,na.rm=T),2)*0.01) %>%
  distinct(Mean_R2,R2_sd,Mean_OOB_Error,Error_sd)

# summarizing VIMP results
Height_VIMP_results <-Height_VIMP %>%
  group_by(Variable) %>%
  summarize(Mean_VIMP = round(mean(Variable_Importance,na.rm=T),2),
            VIMP_sd = round(sd(Variable_Importance,na.rm=T),2)) %>%
  arrange(Mean_VIMP)


```


### Saving Dataframes of Results
```{r Height Results Save, message=FALSE, warning=FALSE}

#show results
Height_results
Height_VIMP_results

#save results
write_csv(Height_results,
          paste0(local_path,"/Data/Final/Height_RF_2023.csv"))

write_csv(Height_VIMP_results,
          paste0(local_path,"/Data/Final/Height_RF_VIMP_2023.csv"))

#clean
rm(list=setdiff(ls(), 
                c("rugosity","local_path","model_data")))

gc()

```


***


## Canopy Rugosity

### Running RF 

```{r Canopy Rugosity RF, message=FALSE, warning=FALSE}

#empty matrix to store error and r2 from model
Rugosity_model<- matrix(nrow=10,ncol=2)

#empty dataframe to store variable importance
Rugosity_VIMP <- data.frame()


#run model 10 times and store results in empty dataframes
for(i in 1:10){

  #----RUNNING MODEL----#
  
  # Run the random forest model with rugosity as response
  # (making sure data is dataframe not tibble)
  RF_Rugosity<- rfsrc(Rugosity~.,
                            data=as.data.frame(rugosity),
                            ntree = 5000, #number of trees
                            importance = "permute") #return VIMP calculated with permutation
  
  #----STORING RESULTS----#
  
  # R2
  Rugosity_r2_i <- round(100* (1-mean(RF_Rugosity$err.rate,na.rm=T)/var(RF_Rugosity$yvar, na.rm = TRUE)),2)
  # error rate
  Rugosity_error_i <- mean(RF_Rugosity$err.rate,na.rm=T)
  # storing
  Rugosity_model[i,] <- c(Rugosity_r2_i,Rugosity_error_i)
  
  #----STORING VIMP----#
  
  # Get variable importance and organize it
  Rugosity_vimp_i <- as.data.frame(RF_Rugosity$importance) %>%
    rownames_to_column(var = "Variable") %>%
    dplyr::rename(Variable_Importance = `RF_Rugosity$importance`) %>%
    arrange(Variable_Importance)
  
  # storing
  Rugosity_VIMP = rbind(Rugosity_VIMP,Rugosity_vimp_i)

}

```


### Summarizing Results

```{r Rugosity Result Summarize, message=FALSE, warning=FALSE}

# turning matrix of results to dataframe
Rugosity_model <- as.data.frame(Rugosity_model)
colnames(Rugosity_model) <- c("R2","Error")

# summarizing
Rugosity_results <- Rugosity_model %>%
  mutate(Mean_R2 = round(mean(R2,na.rm=T),2)*0.01,
         R2_sd = round(sd(R2,na.rm=T),2)*0.01,
         Mean_OOB_Error = round(mean(Error,na.rm=T),2)*0.01,
         Error_sd = round(sd(Error,na.rm=T),2)*0.01) %>%
  distinct(Mean_R2,R2_sd,Mean_OOB_Error,Error_sd)

# summarizing VIMP results
Rugosity_VIMP_results <-Rugosity_VIMP %>%
  group_by(Variable) %>%
  summarize(Mean_VIMP = round(mean(Variable_Importance,na.rm=T),2),
            VIMP_sd = round(sd(Variable_Importance,na.rm=T),2)) %>%
  arrange(Mean_VIMP)

```


### Saving Dataframes of Results

```{r Rugosity Results Save, message=FALSE, warning=FALSE}

#show results
Rugosity_results
Rugosity_VIMP_results

#save results
write_csv(Rugosity_results,
          paste0(local_path,"/Data/Final/Rugosity_RF_2023.csv"))

write_csv(Rugosity_VIMP_results,
          paste0(local_path,"/Data/Final/Rugosity_VIMP_2023.csv"))

#clean
rm(list=ls())

gc()

```