---
title: "Canopy Structure Manuscript Model Interactions"
author: "Roí Ankori-Karlinsky"
date: "5/25/2022"
output:
  html_document: 
    toc: yes
    toc_depth: 5
    toc_float: yes
    theme: journal
    fig_width: 9
    fig_height: 8
---


```{r setup, include=FALSE,echo=FALSE}

# load packages 

#Basic
library(tidyverse) # Grammar for data manipulation
library(parallel) # mclapply for multicore processing

#Spatial analysis
library(sp) #provides object oriented classes for spatial data in R 
library(sf) #similar to sp
library(rgdal) # raster analysis
library(raster) #provides classes and methods for raster datasets
library(spdplyr) #spatial data manipulation

#Data checking
library(naniar) # Visualization of missing values
library(Amelia) # Teting missing columns
library(corrplot) #correlation plots

#Random Forest
library(randomForestSRC) # random forests for survival, regression and
library(pdp) #partial dependence plots
library(randomForest) #random forest
library(randomForestExplainer) #unpacking RF results
library(pre) #Interactions between predictors
library(iml) #Interactions between predictors

# Visualization
library(ggRandomForests) # ggplot2 random forest figures
library(gridExtra) # allows plotting of multiple ggplots in one
library(grid) #same as above
library(colorRamps) # Colors for graphing
library(RColorBrewer) # Colors for graphing
library(cowplot) #putting plots together
library(plot3D) #Surface plots

#Don't include code chunks when outputting markdown
knitr::opts_chunk$set(
echo=FALSE
)

```

```{r data prep, message=FALSE, warning=FALSE, echo=F}

#File path
local_path <- 'C:/Users/roiak/Documents/Damage Project/Rugosity'

full_d <- read_csv(paste0(local_path,"/Data/Final/Full_Data_For_Analyses.csv"))#data file

#cleaning data
full_d <- full_d %>%
  #Fixing variables (factors, making sure precipitation is in mm/year)
  mutate(Soil = factor(Soil),
         Max_Height = as.numeric(Max_Height),
         Forest_Age = factor(Forest_Age),
         MAP = MAP *0.01,
         Wind_Binary = ifelse(Wind_Binary == 0, "Protected",
                              ifelse(Wind_Binary == 1, "Exposed",NA))
         )  %>%
  #Removing outliers and sites below sea-level
  filter(CHM > 2 & CHM <40,
         Elev_Mean >=0)

#Fixing wind-exposure
full_d$Wind_Binary <- factor(full_d$Wind_Binary,
                              levels = c("Protected","Exposed"))

#Creating dataset for each response variable
height <- full_d %>%
  dplyr::select(CHM,
                Forest_Age,
                MAP,
                Wind_Binary,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean) %>%
  na.omit(.)


rugosity <- full_d %>%
  dplyr::select(Rugosity,
                CHM,
                Forest_Age,
                MAP,
                Wind_Binary,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean) %>%
  na.omit(.)

```

# OVERVIEW

## Questions

- **(1)** Does regular exposure to non-hurricane winds reduce canopy structural complexity in tropical forests?  
- **(2)** To what extent is this effect mediated by climate, topography, substrate, forest age and past disturbances?

## Goals

This script checks for Random Forest (RF) model interactions between predictors of:     
      - **(1)** Canopy Height
      - **(2)** Canopy Top Rugosity (sd of height along moving windows) - Vertical variation for top of canopy  
      
      
The predictors are:    
      - Soil type and soil water availability
      - Forest age  
      - Wind exposure (calculated with prevalent directions and using a 30m DEM)
      - Elevation mean and range (30m USGS DEM)
      - Slope (30m USGS DEM)
      - Long term mean annual precipitation 1963-1995 (PRISM)
      - Past exposure to hurricane Hugo in 1989 and hurricane Georges in 1998 (Boose et al., 2004)

## Data

The LiDAR data come from 2016 flights by the USGS, which were then normalized by Lora Murphy of the Cary Institute.
Lora also calculated a canopy height model (CHM) and top rugosity.

Study sites were selected by stratified sampling randomly across forest age,
Then creating a required 250m minimum distance between points,
Then creating 30m buffers around to have a scale that captures a forest stand,
Then filtering out non-forest points if they contain non-forest elements within the buffer.

This resulted in 20,660 study sites across Puerto Rico for the 30m buffer.

(Thanks Dr. Jazlynn Hall with help for the code!)


*** 

# Interactions

Testing for interactions with pre and iml packages
Calculating the H-statistic 
Code adapated from Jaz Hall, thanks!

## Height

```{r Height Interactions, message=FALSE, warning=FALSE, fig.height=8, fig.width=14, dpi=600}

#Doing this on a subsample of the data since it's so large

#random number generator
set.seed(10)
#Random subsample of data
interactions.data<- height[sample(nrow(height),
                                              3000), ]
#removing NAs
interactions.data<- na.omit(interactions.data)

#Running interaction model
interaction_model <- pre::pre(CHM ~ ., 
                              data=interactions.data[complete.cases(interactions.data),],
                              verbose=F)

#Check model diagnostics
summary(interaction_model)

# use the iml package to test interactions 
# Create a model object the package can recognize
mod=Predictor$new(interaction_model,
                  data=interactions.data[-which(names(interactions.data)=="CHM")])

# Measure the interaction strength
ia = Interaction$new(mod)

#make naming better
ia$results <- ia$results %>%
  mutate(.feature = case_when(
    .feature == "CHM" ~ "Canopy Height (m)",
    .feature == "MAP" ~ "Mean Annual Precipitation (mm/yr)",
    .feature == "Forest_Age" ~ "Forest Age",
    .feature == "Wind_Binary" ~ "Chronic Wind-Exposure",
    .feature == "Elev_Mean" ~ "Elevation (m)",
    .feature == "AWS_mean" ~ "Available Soil Water Storage (cm)",
    .feature == "Slope" ~ "Slope (degrees)",
    .feature == "Soil" ~ "Soil Type"
    ))

#save results
write_csv(ia$results,
          paste0(local_path,"/Data/Final/Height Interaction.csv"))

#Remove items and clean space
rm(ia,mod,interaction_model,interactions.data)
gc()

```

***

## Rugosity

```{r Rugosity Interactions, message=FALSE, warning=FALSE, fig.height=8, fig.width=9, dpi=600}

#Doing this on a subsample of the data since it's so large

#random number generator
set.seed(10)
#Random subsample of data
interactions.data<- rugosity[sample(nrow(rugosity),
                                              3000), ]
#removing NAs
interactions.data<- na.omit(interactions.data)

#Running interaction model
interaction_model <- pre::pre(Rugosity ~ ., 
                              data=interactions.data[complete.cases(interactions.data),],
                              verbose=F)

#Check
summary(interaction_model)

# use the iml package to test interactions 
# Create a model object the package can recognize
mod=Predictor$new(interaction_model,
                  data=interactions.data[-which(names(interactions.data)=="Rugosity")])

# Measure the interaction strength
ia = Interaction$new(mod)

#make naming better
ia$results <- ia$results %>%
  mutate(.feature = case_when(
    .feature == "CHM" ~ "Canopy Height (m)",
    .feature == "MAP" ~ "Mean Annual Precipitation (mm/yr)",
    .feature == "Forest_Age" ~ "Forest Age",
    .feature == "Wind_Binary" ~ "Chronic Wind-Exposure",
    .feature == "Elev_Mean" ~ "Elevation (m)",
    .feature == "AWS_mean" ~ "Available Soil Water Storage (cm)",
    .feature == "Slope" ~ "Slope (degrees)",
    .feature == "Soil" ~ "Soil Type"
    ))

#save results
write_csv(ia$results,
          paste0(local_path,"/Data/Final/Rugosity Interaction.csv"))


#Remove items and clean space
rm(list=ls())
gc()
```
