---
title: "Canopy Structure Manuscript RF Analyses"
author: "Roi A.K."
date: "2023-03-13"
output:
  html_document: 
    toc: yes
    toc_depth: 5
    toc_float: yes
    theme: journal
    fig_width: 9
    fig_height: 8
---


```{r setup, include=FALSE,echo=FALSE}

# load packages 

#Basic
library(tidyverse) # Grammar for data manipulation
library(parallel) # mclapply for multicore processing
library(vegan)

#Spatial analysis
library(sp) #provides object oriented classes for spatial data in R 
library(sf) #similar to sp
library(rgdal) # raster analysis
library(raster) #provides classes and methods for raster datasets
library(spdplyr) #spatial data manipulation

#Data checking
library(naniar) # Visualization of missing values
library(Amelia) # Teting missing columns
library(corrplot) #correlation plots

#Random Forest
library(randomForestSRC) # random forests for survival, regression and
library(pdp) #partial dependence plots
library(randomForest) #random forest
library(randomForestExplainer) #unpacking RF results
library(pre) #Interactions between predictors
library(iml) #Interactions between predictors

# Visualization
library(ggRandomForests) # ggplot2 random forest figures
library(gridExtra) # allows plotting of multiple ggplots in one
library(grid) #same as above
library(colorRamps) # Colors for graphing
library(RColorBrewer) # Colors for graphing
library(cowplot) #putting plots together
library(plot3D) #Surface plots

#Making sure R is utilizing all available RAM
memory.limit(size=50000)

#File path
local_path <- 'C:/Users/roiak/Documents/Damage Project/Rugosity'

#Don't include code chunks when outputting markdown
knitr::opts_chunk$set(
  echo=FALSE,
  dev = 'png',
  dpi=600,
  fig.path = (paste0(local_path,"/Figures/Canopy Structure Manuscript RF Mar 2023/")
)
)

```

```{r data prep, message=FALSE, warning=FALSE, echo=F}


#load data
full_d <- read_csv(paste0(local_path,"/Data/Final/Full_Data_For_Analyses.csv"))#data file

#cleaning data
full_d <- full_d %>%
  filter(Soil!="Water") %>%
  #Fixing variables (factors, making sure precipitation is in mm/year)
  mutate(Soil = factor(Soil),
         Max_Height = as.numeric(Max_Height),
         Forest_Age = factor(Forest_Age),
         MAP = MAP *0.01,
         Wind_Binary = ifelse(Wind_Binary == 0, "Protected",
                              ifelse(Wind_Binary == 1, "Exposed",NA)),
         Wind_Exposure = round(Wind_Exposure,2),
         Wind_sd = round(Wind_sd,2)
         )  %>%
  #Removing outliers and sites below sea-level
  filter(CHM > 2 & CHM <40, #getting rid of canopy height outliers
         Elev_Mean >=0)

#Creating dataset for each response variable
height <- full_d %>%
  dplyr::select(CHM,
                Forest_Age,
                MAP,
                Wind_Exposure,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean,
                Hugo_exposure) %>%
  na.omit(.)


rugosity <- full_d %>%
  dplyr::select(Rugosity,
                CHM,
                Forest_Age,
                MAP,
                Wind_Exposure,
                Elev_Mean,
                Slope,
                Soil,
                AWS_mean,
                Hugo_exposure) %>%
  na.omit(.)

```

# OVERVIEW

### Questions

- **(1)** Does regular exposure to non-hurricane winds reduce canopy structural complexity in tropical forests?  
- **(2)** To what extent is this effect mediated by climate, topography, substrate, forest age and past disturbances?

### Goals

This script tests whether chronic wind exposure reduces canopy structural complexity across Puerto Rico.
It does this by running 2 Random-Forest (RF) models for two airborne LiDAR-derived canopy metrics:

- **(1)** Canopy Height
- **(2)** Canopy Top Rugosity (sd of height along moving windows) - Vertical variation for top of canopy  

### Data

The LiDAR data come from 2016 flights by the USGS, which were then normalized by Lora Murphy of the Cary Institute.
Lora also calculated a canopy height model (CHM) and top rugosity.

Study sites were selected by stratified sampling randomly across forest age,
Then creating a required 250m minimum distance between points,
Then creating 30m buffers around to have a scale that captures a forest stand,
Then filtering out non-forest points if they contain non-forest elements within the buffer.

This resulted in 20,660 study sites across Puerto Rico for the 30m buffer.

The independent variables used are:
  - Soil type and soil water availability
  - Forest age  
  - Wind exposure (calculated with prevalent directions and using a 30m DEM)
  - Elevation mean and range (30m USGS DEM)
  - Slope (30m USGS DEM)
  - Long term mean annual precipitation 1963-1995 (PRISM)
  - Past exposure to hurricane Hugo in 1989 and hurricane Georges in 1998 (Boose et al., 2004)

***

# Random Forest Models

Here I run 10 RF models for each response variable.
Then I summarize variable importance of predictors as the mean and standard error from all 10 models.
I also report the model error and R$^2$ 


## Canopy Height


### Running RF 

```{r Canopy Height RF, message=FALSE, warning=FALSE}

#running height RF
Height_RF <- randomForest(CHM ~ .,
                          data=height,
                          ntree = 500,
                          importance = T)

Height_RF

```


### Variable importance

```{r Height VIMP, message=FALSE, warning=FALSE}

#variable importance plot
varImpPlot(Height_RF) 

```

### Predictions


```{r Height explain, message=FALSE, warning=FALSE}

explain_forest(Height_RF, interactions = TRUE, data = height)

gc()

```

```{r clean height}
rm(Height_RF,height)
gc()

```




***


## Canopy Rugosity

### Running RF 

```{r Canopy Rugosity RF, message=FALSE, warning=FALSE}

#running rugosity RF
Rugosity_RF <- randomForest(Rugosity ~ .,
                          data=rugosity,
                          ntree = 500,
                          importance = T)

Rugosity_RF
gc()

```


### Variable importance

```{r Rugosity VIMP, message=FALSE, warning=FALSE}

#variable importance plot
varImpPlot(Rugosity_RF) 
gc()

```

### Predictions


```{r Rugosity explain, message=FALSE, warning=FALSE}

explain_forest(Rugosity_RF, interactions = TRUE, data = rugosity)

```

